### 2022-04-22 greta/sn_mcmc.Rmd
sn mcmc :turtle:

set.seed(4242)
train_reps <- 100
train_t_max <- 250
test_t_max <- 250
test_reps <- 100

@@

```{r}
# View the bistable curve: 
#curve( (0.75 - x)**2 * (0.25 - x), 0.749999, 0.750001)
#curve(0*x,0,1,add = TRUE)

#curve((0.25 - x)**2 * (0.75 - x), 0, 1)
#curve(0*x,0,1,add = TRUE)
```

```{r}
step <- function(N, eta,  a = 0.75, b = 0.25) {
  N <- N + (a - N)**2 * (b - N) + eta
  N <- np.clip(N, 0, 1)
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 1250L,
                       p = list(
                       sigma = 1e-2,
                       a = 0.75, 
                       b = 0.25)) {
  eta <- rnorm(t_max, 0, p$sigma) # mu = 0, no drift
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], a = p$a, b = p$b)
  }
  tibble::tibble(t = 1:t_max, N = N)
}

train <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=train_t_max), .id = "i")
test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

## show training data
train |> ggplot(aes(t, N, group=i)) + geom_line(alpha=0.2)
```

```{r}
gsims <- train |> group_by(i) |> mutate(xt1 = lead(N)) |> filter(t<max(t))

x_t <- gsims$N
x_t1 <- gsims$xt1

library(greta)
a <- uniform(0, 10)
b <- uniform(0, 10)
sigma <- uniform(0, 10)
mean <-  x_t + (a - x_t) ^ 2 * (b - x_t)
distribution(x_t1) <- normal(mean, sigma )
m <- model(a, b, sigma)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("mcmc_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 10000, warmup = 5000, chains = 4, verbose = FALSE)
})
```

@@

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps)) + 
(ggplot(rep_scores) + geom_histogram(aes(logs)) )

>>>>

set.seed(4242)
train_reps <- 1
train_t_max <- 75
test_t_max <- 75
test_reps <- 100

@@

```{r}
step <- function(N, 
                 eta,
                 t,
                 pars) {

        h <- np.clip(pars$h_init + pars$alpha * t, 0, 0.27)
        N <- N + pars$r * N * (1 - N / pars$K) - 
          h * (N**2 / (pars$s**2 + N**2)) + eta
        N <- np.clip(N, 0, 1)
        N
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 250L,
                     pars = list(
                        r = 1,
                        K = 1,
                        s = 0.1,
                        h_init = .15,
                        alpha = 0.0015,
                        mu = 0,
                        sigma = 0.00
                      )
                    ) {
  eta <- rnorm(t_max, pars$mu, pars$sigma) # mu = 0, no drift
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], t, pars)
  }
  tibble::tibble(t = 1:t_max, N = N)
}
```

```{r}
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02
          )
t_max <- train_t_max + test_t_max
sims <- purrr::map_dfr(1:train_reps, 
                       \(i) simulate(t_max=t_max, pars = pars), 
                       .id = "i")

train <- sims |> filter(t <= train_t_max)
test <- sims |> filter(t > train_t_max)
```

```{r}
sims <- bind_rows(test, train, .id = "data")
sims |> 
  ggplot(aes(t, N, group=interaction(i,data), col=data)) + 
  geom_line()
```

```{r}
gsims <- train |> group_by(i) |> mutate(xt1 = lead(N)) |> filter(t<max(t))
x_t <- gsims$N
x_t1 <- gsims$xt1
t <- gsims$t

library(greta)
r <- uniform(0, 10)
K <- uniform(0, 10)
s <- uniform(0,10)
h_init <- uniform(0,10)
alpha <- uniform(0, 10)
sigma <- uniform(0, 10)
h <- h_init + alpha * t
mean <-  x_t + r*x_t*(1 - x_t / K) - h*(x_t^2 / (s^2 + x_t^2))
distribution(x_t1) <- normal(mean, sigma)
m <- model(r, K, s, h_init, alpha, sigma)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("sn_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 60000, warmup = 50000, chains = 4, verbose = FALSE)
})
```

@@

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps), bins=20) + 
(ggplot(rep_scores) + geom_histogram(aes(logs), bins=20) )
